---
title: "junction_exploration_1"
output: html_document
---

```{r}
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(ggplot2)
```

Change this to the base directory of where you downloaded the Snaptron data.  Use the `get.sh` script to obtain the Snaptron data:

```{r setup}
# pre-computed
knitr::opts_knit$set(root.dir = '~/snaptron_data')
```

Read a Snaptron-style metadata file.

```{r read_metadata}
# pre-computed
read_metadata <- function(project) {
    fread(file.path(project, 'samples.tsv'), sep='\t', header=T, quote="")
}
```

Read a Snaptron-style junction-evidence matrix, while removing unnecessary columns.

```{r read_junction_annotation}
# pre-computed
read_junction_annotation <- function(project) {
    col_select <- c(5, 7, 10, 11, 13, 14, 15, 16)
    colnms <- scan(file.path(project, 'junctions.header.tsv'), what='character')
    colnms <- colnms[col_select]
    jx_fn <- file.path(project, 'junctions.bgz')
    cmd <- paste0(
        "gzip -dc ", jx_fn, " | cut -f",
        paste(col_select, collapse=','))
    jxs <- fread(cmd, sep='\t', header=F)
    colnames(jxs) <- colnms
    jxs
}
```

Grab the junctions for the set of >50K samples from SRA.

```{r jx_read_srav2_data}
# pre-computed
jx_srav2 <- read_junction_annotation('srav2')
```

```{r jx_read_srav2_metadata}
meta_srav2 <- read_metadata('srav2')
```

```{r size}
print(paste('# junctions:', nrow(jx_srav2)))
print(paste('# samples:', nrow(meta_srav2)))
```

Prepare the data for the cumulative-annotated plot shown in Fig 2a of the [Intropolis paper](https://doi.org/10.1186/s13059-016-1118-6).  Our plot will be different as we're using the (bigger) SRAv2 dataset and because we're only considering the GENCODE annotation.

```{r cumulative_plot}
# pre-computed
prepare_cumulative <- function(jxs, ann='gC38') {
    jxs %>% group_by(samples_count) %>%
        summarise(
            neither  = sum(!grepl(ann, left_annotated) &  !grepl(ann, right_annotated)),
            one      = sum( grepl(ann, left_annotated) !=  grepl(ann, right_annotated)),
            both_un  = sum( grepl(ann, left_annotated) &   grepl(ann, right_annotated) & annotated == 0),
            both_ann = sum(annotated == 1),
            n        = n()) %>%
        mutate(
            ann_cum = both_ann,       both_cum = ann_cum + both_un,
            one_cum = both_cum + one, neither_cum  = one_cum + neither) %>%
        select(samples_count, ann_cum, both_cum, one_cum, neither_cum, n) %>%
        arrange(desc(samples_count)) %>%
        mutate(
            ann_cum=cumsum(ann_cum), both_cum=cumsum(both_cum),
            one_cum=cumsum(one_cum), neither_cum=cumsum(neither_cum)) %>%
        gather(category, count, ann_cum, both_cum, one_cum, neither_cum)
}
```

```{r jx_srav2_cumulative}
# pre-computed
jx_srav2_cumulative <- prepare_cumulative(jx_srav2)
```

Make the cumulative-annotated plot.

```{r}
cumulative_plot <- function(jx_cumulative) {
    jx_cumulative$category <- plyr::mapvalues(jx_cumulative$category,
        c('ann_cum', 'both_cum', 'neither_cum', 'one_cum'),
        c('Annotated', 'Exon Skip', 'Neither Annotated', 'One Annotated'))
    ggplot(jx_cumulative, aes(x=samples_count, y=count, color=category)) +
        geom_line() +
        coord_cartesian(xlim=c(0, max(jx_cumulative$samples_count)*0.5), ylim=c(0, 750000)) +
        labs(x='Jxs appearing in at least this # samples', y='Count') +
        theme_bw()
}
```

```{r cum_plot_srav2}
cumulative_plot(jx_srav2_cumulative)
```

Because the cumulative-annotated plot makes it hard to discern particular percentages, we summarize some in a table.

```{r annotation_table}
annotation_table <- function(jxs, cutoffs=c(0.01, 0.02, 0.05, 0.10, 0.20, 0.33, 0.5, 0.75)) {
    counts <- sapply(cutoffs, function(x) { ceiling(max(jxs$samples_count)*x) } )
    jx_cutoffs <- jxs %>%
        filter(samples_count %in% counts) %>%
        spread(category, count) %>%
        mutate(
            ann_pct=ann_cum*100.0/neither_cum,
            both_pct=(both_cum-ann_cum)*100.0/neither_cum,
            one_pct=(one_cum-both_cum)*100.0/neither_cum,
            neither_pct=(neither_cum-both_cum)*100.0/neither_cum) %>%
        select(ann_pct, both_pct, one_pct, neither_pct)
    jx_cutoffs$pcts <- cutoffs * 100
    colnames(jx_cutoffs) <- c(
        'Annotated', 'ExonSkip', 'OneAnnotated',
        'NeitherAnnotated', 'PercentOfSamples')
    jx_cutoffs
}
```

```{r annotation_table_srav2}
annotation_table(jx_srav2_cumulative)
```
